Solving the power flow problem on integrated transmission-distribution networks: a review and numerical assessment

### 2020-10-28 IEEEtran/paper.tex
Draft_before_revision

% packages needed for plotting results
\usepackage{tikz}
\usetikzlibrary{trees}
\usepackage{pgfplots}
%\pgfplotsset{compat=1.15}
\usepackage{multirow}
\usepackage{booktabs} %for fancy tables
\usepackage{adjustbox}

\usepackage{todonotes} 
% \usepackage[disable]{todonotes} %Hides all(!) todonotes
% todonotes appearance
\newcommand{\addref}{\todo[color=red!40,caption={}]{Add reference.}}
\newcommand{\rewrite}[1]{\todo[inline,color=yellow!40,caption={}]{#1}} % the caption={} is needed if you want to include itemize and such in the todonotes, see: https://tex.stackexchange.com/questions/54064/itemize-inside-todonote
\newcommand{\question}[1]{\todo[inline,color=blue!40,caption={}]{#1}}
\newcommand{\includes}[1]{\todo[inline,color=black!10,caption={}]{#1}}
% % Don't print any [non-inline] notes (will still show up in list of todo notes, to disable this, see: http://tex.stackexchange.com/questions/287394/todonotes-show-inline-hide-all-others)
% \makeatletter
% \renewcommand{\@todonotes@drawMarginNoteWithLine}{}
% \makeatother

@@

Power flow computations are used to simulate the transport, generation, and consumption of power in electricity networks. The simulations are important for safe operation and planning of the electricity grid. An electricity grid consists of one transmission network, responsible for transport of power over large distances, and several distribution networks, responsible for the transport to end-consumers. Transmission networks transport high-voltage power to substations from where power is converted to a lower voltage level. The distribution network transports low-voltage power from substations to end-consumers. These transmission and distribution domains are separately analyzed where one domain is modeled in detail and the other is simplified in this model to one bus. Eg: in the analysis of the transmission system, the distribution network is simplified to one load in the transmission model. In the analysis of the distribution system,  the transmission network is simplified as a generator.  \newline\newline
The power system is changing: more renewable energy is entering the grid at distribution level, we see an increase of demand-side participation as a mechanism to balance frequency, and we see an increase of electricity consumption. The changing environment requires a more detailed analysis of the electricity network and a more complex model of the interaction between the networks. Integrated transmission-distribution network models study the interaction that these networks have on each other. But it is not straight-forward to integrate these separate domains. The networks have different characteristics what has resulted in different network models that cannot be easily integrated. The most important difference is that the transmission network is balanced and therefore modeled in single-phase while the distribution network is in general not balanced and modeled in three-phase. 
Furthermore, the different characteristics have lead to the development of different algorithms to solve the power flow problem. The power flow problem is a nonlinear problem which is solved using iterative methods. Well-known algorithms to solve transmission systems are: Gauss-Seidel, Newton-Raphson, Decoupled loadflow, and DC loadflow \cite{Schavemaker2008chap6}. These solvers have been adapted to distribution solvers, among which the most well-known are: Newton-Raphson current mismatch \cite{Garcia2000}\cite{Abdel-akher2005}, implicit $Z_{bus}$ \cite{Teng2002a}\cite{Chen1991a}, BIBC/BCBV \cite{Teng2008}\cite{Farag2011}, and Forward/Backward Sweep \cite{Moghaddas2009}\cite{Eminoglu2008} methods. Besides the technical complexity of integrating network models, we have the issue of confidentiality. System operators are not willing or not allowed to share network information with other system operators. \textit{One can reason that conventional models will at a certain moment not suffice anymore. \todo{De volgende zin moet er helemaal niet in of vriendelijker verwoord worden}Lagging legislation will then follow the necessary change towards integrating network models. 
}\newline\newline 
Integrated networks can be modeled as homogeneous and as hybrid networks. A homogeneous network model is a three-phase representation of both the transmission and distribution network. A hybrid network model consists of a single-phase transmission and a three-phase distribution network. Several numerical methods have been developed to solve integrated networks. They can be classified as a unified \cite{Taranto2008} or as a splitting method \cite{Sun2005}.  The unified methods solve the integrated system as a whole. The splitting method divides the networks into a master and a slave and defines an extra iterative scheme between them. At every iteration, the separated systems are solved on its own. The integrated system is solved once convergence on the boundary between the systems have been reached. This article reviews the proposed integration methods and analyzes them by running simulations on small test-cases. We pay special attention to their convergence behaviour: CPU-time and the number of iterations.

@@

\subsection{Modeling elements in three-phase configuration}\hl{Voeg hier een opmerking toe dat ene noads elements zijn, andere series elements. dat we total Ybus matrix moeten bepalen en daar series elementen belangrijk zijn.}
Loads, shunts, and transformers in three-phase networks are placed in a Wye or Delta configuration. Other elements that we have not seen so far in transmission networks are regulators. Regulators are placed along the distribution feeder to regulate the voltage distribution. 

@@

Loads are modeled as buses where active and reactive power ($P$ and $Q$) are specified. They are modeled as a function of the voltage in which the load-model type (constant power/current/impedance) determines the relation between power and voltage:  

@@

@@

    \hl{quote: A constant current load is one which varies its internal resistance to achieve a constant current regardless of the voltage which is being fed to it and therefore the power will vary. In other words, a constant current load would draw approx the same amount of current, despite wide variations in the source voltage or other impedances in the circuit.}

@@

@@

Loads are physically connected in a grounded Wye or ungrounded Delta configuration. In the grounded Wye configuration all the three phases are connected to a single neutral point. These neutral point is connected to the ground by a fourth wire: the neutral conductor. The voltage is specified line-to-neutral.
In an ungrounded Delta configuration the loads are connected phase-to-phase without a neutral conductor. Here, the voltage is specified phase-to-phase, also called line-to-line. The line-to-line voltages of a three-phase Delta-configuration are: $V^{ab}$, $V^{bc}$, and $V^{ca}$.  The line-to-line voltages and line-to-neutral voltages ($V^a$, $V^b$, and $V^c$) in this configuration are related as follows:

@@

We solve the three-phase distribution systems with the NR-TCIM method, which implies that we need the specified currents $I_s$ in our mismatch equation \hl{(add reference to equation)} instead of specified powers. As currents are never measured in a power system, they are conducted from the specified power according the relationship: 

@@

If we divide the expressions of the loads $S$ by $V$ and take the conjugate of this new relation we get the expression for the specified currents:

@@

\subparagraph{Wye - Z }
\begin{equation}
    I := (\frac{S|V|^2}{V})^* = ({P_0+\iota Q_0})^*\frac{|V|^2}{V^*} = ({P_0+\iota Q_0})^*\frac{VV^*}{V^*} =  ({P_0+\iota Q_0})^*V
\end{equation}
Or in three-phase matrix: 
\begin{align}
\begin{bmatrix}
    I^a\\
    I^b\\
    I^c
    \end{bmatrix}_i = \begin{bmatrix}
   ({ P^a +\iota Q^a})^*{V^a}\\
    ({ P^b +\iota Q^b})^*{V^b}\\
    ({ P^c +\iota Q^c})^*{V^c}
    \end{bmatrix}_i 
    \end{align}
\paragraph{Delta loads}
Delta loads are expressed by line-to-line voltages and currents while we need line-to-neutral currents for our calculations. We express the line-to-neutral currents directly for the three load models using \begin{align}
    I^a=I^{ab}-I^{ca},\\
    I^b=I^{bc}-I^{ab},\\
    I^c=I^{ca}-I^{bc}.
\end{align}
\subparagraph{Delta - P}
\begin{align}
\begin{bmatrix}
    I^a\\
    I^b\\
    I^c
    \end{bmatrix}_i := \begin{bmatrix}
    I^{ab}-I^{ca}\\
    I^{bc}-I^{ab}\\
    I^{ca}-I^{bc}
    \end{bmatrix}_i = \begin{bmatrix}
   (\frac{ P^{ab} +\iota Q^{ab}}{V^a-V^b})^* - (\frac{ P^{ca} +\iota Q^{ca}}{V^c-V^a})^* \\
    (\frac{ P^{bc} +\iota Q^{bc}}{V^b-V^c})^* - (\frac{ P^{ab} +\iota Q^{ab}}{V^a-V^b})^*\\
    (\frac{ P^{ca} +\iota Q^{ca}}{V^c-V^a})^* - (\frac{ P^{bc} +\iota Q^{bc}}{V^b-V^c})^*
    \end{bmatrix}_i 
    \end{align}
\subparagraph{Delta - I }
\begin{align}
\begin{bmatrix}
    I^a\\
    I^b\\
    I^c
    \end{bmatrix}_i = \begin{bmatrix}
    ({ P^{ab} +\iota Q^{ab}})^*\frac{(V^a-V^b)}{|V^a-V^b|} - ({ P^{ca} +\iota Q^{ca}})^*\frac{(V^c-V^a)}{|V^c-V^a|}  \\
    ({ P^{bc} +\iota Q^{bc}})^*\frac{(V^b-V^c)}{|V^b-V^c|} - ({ P^{ab} +\iota Q^{ab}})^*\frac{(V^a-V^b)}{|V^a-V^b|}\\
    ({ P^{ca} +\iota Q^{ca}})^*\frac{(V^c-V^a)}{|V^c-V^a|} -  ({ P^{bc} +\iota Q^{bc}})^*\frac{(V^b-V^c)}{|V^b-V^c|}
    \end{bmatrix}_i 
    \end{align}

@@

\bottomrule
\end{tabular}
\end{adjustbox}
\end{table}
\hl{Note: update Jacobian according new load models}

\subsubsection{Transformers}
Transformers are placed between two networks that have different voltage levels. The primary side ($p$) of the transformer is connected to the higher voltage network and the secondary side ($s$) is connected to the voltage level. The transformer in a power system is represented by two blocks:
\begin{itemize}
    \item An admittance matrix block, $Y^{abc}_T$, consisting of per-unit leakage admittance $y_t$, and  
    \item A shunt block that models core losses. The shunt block can be represented by voltage dependent loads and follow the representation in Tab. \ref{tab:ZIPmodels}. 
\end{itemize}
The admittance matrix block is given in a similar manner as the $Y_{bus}$ matrix of a distribution cable, now consisting of leakage admittance of the transformer: \begin{equation}
    Y^{abc}_T = \begin{bmatrix}
    Y_{pp}^{abc} & Y^{abc}_{ps} \\
     Y_{sp}^{abc} & Y^{abc}_{ss}
    \end{bmatrix}.
\end{equation}
Each block $Y^{abc}_{ij}$, $i,\ j\in\{p,s\}$ has a different configuration. This configuration depends on the connection of the transformer and is either one of the three following blocks: \begin{equation}\resizebox{0.91\columnwidth}{!}{%
$
    Y_1=\begin{bmatrix}
    y_t & 0 & 0 \\
    0 & y_t & 0 \\
    0 & 0 & y_t
    \end{bmatrix},\quad Y_2=\frac{1}{3}\begin{bmatrix}
    2y_t & -y_t & -y_t \\
    -y_t & 2y_t & -y_t \\
    -y_t & -y_t & 2y_t
    \end{bmatrix},\quad\mbox{and}\quad Y_3=\frac{1}{\sqrt{3}}\begin{bmatrix}
    -y_t & y_t & 0 \\
    0 & -y_t & y_t \\
    y_t & 0 & -y_t 
    \end{bmatrix}. $%
    }
\end{equation}
Tab. represents the corresponding block per transformer configuration. \\
The pu leakage admittance $y_t$ consists of resistance $r$ and reactance $x$ of the transformer and should \hl{be corrected for the correct base grounds}, resulting in: 
\begin{equation}
    y_t = \frac{1}{z_t},\quad\mbox{where } z_t = (r+\iota x)(\frac{V_{low}}{V_{base}})^2\frac{S_{base}}{VA},
\end{equation}
\hl{where $r$ and $x$ are given in percentages, $V_{low}$ is the voltage at the secondary level of the transformer, and $VA$ is volt-ampere, the unity of apparent power, of the transformer.  }
Several transformer connections exist, which define the choice of the four blocks of the admittance matrix $Y_T^{abc}$. The correct blocks are described in Tab. \ref{tab:trafoblocks}.

@@

Step-Voltage Regulators (SVRs) are installed along distribution feeders, often after the substation, but also at other locations, to control the voltage along the feeder and to keep it within an acceptable range.  \\
In short, the SVR is an automatic transformer that can change its tap ratio. The tap ratio of a transformer is responsible for the voltage level at the secondary side of a transformer, the output voltage. In a SVR, this voltage is measured; if it falls outside the safety range, the tap ratio changes to a different position to adjust the output voltage. This continues until the voltage is within acceptable limits.
In our load-flow  computations, we are not implementing a SVR that changes its tap ratio automatically. We pre-determine the `safe' tap-ratios based on output data of openDSS. With these data we model the SVR directly. \newline
A SVR is modeled as a block that is placed between any connected pair of buses ($n$,$m$). The  voltages and currents of the primary (bus $n$) and secondary side (bus $m$) of the SVR are related via matrices:  $A_V$, $A_I$, and $Z_R\in\mathbb{C}^{6x6}$. The first one represents voltage gain, the second current gain, and the last impedance. $A_I$ and $A_V$ consists of effective regulator ratio $a_R$ and $Z_R$ consists of pu series impedance $z_R$ of the SVR. Having determined the tap-ratio, $a_R$ is:

@@

    a_R = 1\mp 0.000625tap. 
\end{equation}Three-phase SVRs are commonly configured as Wye, closed-delta, and open-delta. Just like transformers, these configurations determine the entities of the three blocks $A_I$, $A_V$, and $Z_R$. It holds that:  \begin{equation}

@@

With these three blocks we form the SVR-admittance matrix, $Y^{abc}_{R}$, that can be incorporated in the final nodal admittance matrix of the feeder, $Y_{bus}$. \hl{Say this also in transformer section}. 
The admittance matrix $Y^{abc}_R$ looks as follows:

@@

A_I F_R^{-1}Y^{abc}_{nn}A_I^T & -A_I F_R^{-1}Y^{abc}_{nm} \\ 

@@

$Y^{abc}_{ij},\ i,j\in\{n,m\}$ consists of the admittance matrix of the distribution line along which the regulator is placed and $F_R$ and $F_R^{-T}$ are defined as follows: 

@@

    F_R &= \mathbf{I}_{3\times 3}+Y^{abc}_{nn}A_I^TZ_RA_I,\\

@@

If the auto-transformer of the SVR is ideal, it means that the tap-ratio is equal to $0$ and that the impedance matrix is $0$, $Z_R=0$. This results into $F_R$ being the identity matrix: $\mathbf{I}_{3\times 3}$. The admittance matrix of an ideal SVR becomes the following:

@@

A_I Y^{abc}_{nn}A_I^T & -A_I  Y^{abc}_{nm} \\ 

@@

We refer to paper \cite{} on how to derive the blocks $A_I$, $A_V$, and $Z_R$. We list the entities of these blocks, per SVR configuration, in Tab. \ref{tab:svrent}.

@@

Transmission networks and distribution networks are connected to each other via a substation. In separate network models, this substation is modeled as a single-bus containing the information of the other system. In transmission network models, this substation is modeled as a load bus. In distribution network models, the substation is modeled as the slack bus. The total load in the transmission model is equal to the total generated power of the slack bus in the distribution model. The difference of the substation as load bus, is that the parameters $S$ and $V$ are modeled as single-phase quantities, while in the distribution model as three-phase quantities. Figure \ref{fig:boundary} shows this information mismatch. Table \ref{tab:substationbus} shows the single-phase and three-phase representation of $S$, $V$ and $Y$ in the network models. 
\begin{figure}

@@

The literature suggests two approaches to run computations on integrated networks: a unified and a splitting approach, and it suggests two ways of modeling integrated networks: as a homogeneous or as a hybrid network. Two integrated three-phase networks form a homogeneous network\footnote{two single-phase networks can also make a homogeneous network, but we regard only unbalanced three-phase distribution networks.}: both the transmission and distribution network are modeled in three phases. A hybrid network consists of a single-phase transmission network and a three-phase distribution network. The unified methods first integrates the two network into one system and then solves the entire system at once. This has the advantage that it requires only one iterative process to solve the system, but also implies that the system has to be solved with either NR-power or NR-TCIM\footnote{In het geval van dit paper dan}, being disadvantageous for a part of the system. The splitting approach appoints the transmission network as the master and the distribution network as the slave. It iterates between the two networks and at each iteration, it solves the networks separately. The master-slave splitting uses an extra iterative scheme besides the transmission and distribution schemes, but this has the advantage that the separate systems can be solved with the preferred Newton-Raphson algorithm. We call the unified approach applied to homogeneous networks the full three-phase (F3P) method and applied to hybrid networks the interconnected (IC) method. We call all splitting methods Master-Slave splitting (MSS) methods, as described in the literature \cite{Sun2008}, and define them based on their option, eg: the splitting approach applied to hybrid networks is called the MSS-hybrid method. Figure \ref{fig:algorithms} gives an overview of the methods. 

@@

>>>>

% packages needed for plotting results
\usepackage{tikz}
\usetikzlibrary{trees}
\usepackage{pgfplots}
%\pgfplotsset{compat=1.15}
\usepackage{multirow}
\usepackage{booktabs} %for fancy tables
\usepackage{adjustbox}
\usepackage{multicol}
\usepackage{todonotes} 
% \usepackage[disable]{todonotes} %Hides all(!) todonotes
% todonotes appearance
\newcommand{\addref}{\todo[color=red!40,caption={}]{Add reference.}}
\newcommand{\rewrite}[1]{\todo[inline,color=yellow!40,caption={}]{#1}} % the caption={} is needed if you want to include itemize and such in the todonotes, see: https://tex.stackexchange.com/questions/54064/itemize-inside-todonote
\newcommand{\question}[1]{\todo[inline,color=blue!40,caption={}]{#1}}
\newcommand{\includes}[1]{\todo[inline,color=black!10,caption={}]{#1}}
% % Don't print any [non-inline] notes (will still show up in list of todo notes, to disable this, see: http://tex.stackexchange.com/questions/287394/todonotes-show-inline-hide-all-others)
% \makeatletter
% \renewcommand{\@todonotes@drawMarginNoteWithLine}{}
% \makeatother

@@

Power flow computations are used to simulate the transport, generation, and consumption of power in electricity networks. The simulations are important for safe operation and planning of the electricity grid \todo{source}. A national electricity grid consists of one transmission network, responsible for transport of power over large distances, and several distribution networks, responsible for the transport to end-consumers. Transmission networks transport high-voltage power to substations from where power is converted to a lower voltage level. The distribution network transports low-voltage power from substations to end-consumers. These transmission and distribution domains are currently separately analyzed: In the analysis of the first, the transmission network is modeled in detail while the distribution network is modeled as an equivalent load. In the analysis of the distribution system, this system is modeled in detail while the transmission network is simplified as a power supply source.  \newline\newline
The power system is changing: 1) more renewable energy is entering the grid at distribution level, 2) we see an increase of demand-side participation as a mechanism to balance frequency, and 3) we see an increase of electricity consumption\todo{source}. The changing environment requires a more detailed analysis of the electricity network and a more complex model of the interaction between the networks. Integrated transmission-distribution network models study the interaction that these networks have on each other. But it is not straight-forward to integrate these separate domains. The networks have different characteristics what has resulted in different network models that cannot be easily integrated. The most important difference is that the transmission network is a balanced network: Power is generated in three phases but voltage and current of other phases have the same magnitude and are equally shifted in phase by $2/3\pi$ rad. Therefore, the transmission network is modeled in single-phase while the distribution network is in general not balanced and thus modeled in three-phase \todo{source}. 
Furthermore, the different characteristics have lead to the development of different algorithms to solve the power flow problem of different systems. The power flow problem is a nonlinear problem which is solved using iterative methods. Well-known algorithms to solve transmission systems are: Gauss-Seidel, Newton-Raphson, Decoupled loadflow, and DC loadflow \cite{Schavemaker2008chap6}. These solvers have been adapted to distribution solvers, among which the most well-known are: Newton-Raphson current mismatch \cite{Garcia2000}\cite{Abdel-akher2005}, implicit $Z_{bus}$ \cite{Teng2002a}\cite{Chen1991a}, BIBC/BCBV \cite{Teng2008}\cite{Farag2011}, and Forward/Backward Sweep \cite{Moghaddas2009}\cite{Eminoglu2008} methods. Besides the technical complexity of integrating network models, we have the issue of confidentiality. System operators, the instances responsible for safe operation of their electricity grid,  are not willing or not allowed to share network information with other system operators\todo{source}. One can reason that the ongoing energy transition and electrification is going to push the limits of conventional models until they do not suffice anymore.  It is probable that legislation and current behaviour will then follow the necessary change towards the use of integrated network models. 
\newline\newline 
The study of integrated network models is an emerging field. A literature review shows that multiple approaches exist that can roughly be categorized into 1) co-simulation frameworks that combine multiple domains that can be studied using different \textit{formats} and 2) stand-alone analysis frameworks that study integrated network models using the same \textit{software tools} \cite{https://ieeexplore.ieee.org/stamp/stamp.jsp?tp=&arnumber=8585926}. Co-simulation frameworks are modular and well-suited for HPC architectures which make them ideal for simultaneous analysis of large transmission and distribution networks (100s to 1000s of buses) \cite{https://www.nrel.gov/docs/fy16osti/65550.pdf}. Stand-alone frameworks have been depreciated by the co-simulation community because of the high costs of simulation \cite{https://ieeexplore.ieee.org/stamp/stamp.jsp?tp=&arnumber=8585926}. While efficiently integrated stand-alone models require significantly less communication and should therefore be compatible with co-simulation techniques. Literature shows that studies focusing on stand-alone models have \textit{thus far} resulted in the creation of different modeling techniques. These techniques have been tested on small-size transmission and distribution networks in order to check their feasibility \cite{Taranto2008}\cite{Sun2015}. In this work, we give an insight in the performance of the stand-alone methods and their ability to be efficiently scaled to large-size networks: We compare the existing stand-alone methods on convergence and CPU-time by applying them on several transmission-distribution networks. The contributions of this work can be summarized as follows: 
    \item An extensive description of existing stand-alone methods that have been categorized according their modeling and integration approach. Network models can be divided into hybrid and homogeneous network models. A homogeneous network model is a three-phase representation of both the transmission and distribution network. A hybrid network model consists of a single-phase transmission and a three-phase distribution network. Stand-alone integration approaches can be classified as a unified \cite{Taranto2008} or as a splitting method \cite{Sun2005}. The unified method  solves the integrated system as a whole. The splitting method is similar to the co-simulation approach where the separate domains are solved on its own and coupled using an iterative approach;
    \item A comparison of these methods using Matlab \todo{source} in which several test-cases are created to simulate the convergence behavior, where special attention is paid to CPU-time and number of iterations;
    \item An analysis of convergence characteristics under varying circumstances such as a high level of distributed generation and the integration of multiple distribution networks into one transmission-network;
    \item An insight into how these methods should adapt according to HPC applications in order to solve realistically sized networks (up to 1000000s buses). 

% Integrated networks can be modeled as homogeneous and as hybrid networks. A homogeneous network model is a three-phase representation of both the transmission and distribution network. A hybrid network model consists of a single-phase transmission and a three-phase distribution network. Several numerical methods have been developed to solve integrated networks. They can be classified as a unified \cite{Taranto2008} or as a splitting method \cite{Sun2005}.  The unified methods solve the integrated system as a whole. The splitting method divides the networks into a manager and a fellow and defines an extra iterative scheme between them. At every iteration, the separated systems are solved on its own. The integrated system is solved once convergence on the boundary between the systems have been reached. This article reviews the proposed integration methods and analyzes them by running simulations on small test-cases. We pay special attention to their convergence behaviour: CPU-time and the number of iterations.

@@

\subsection{Modeling elements in three-phase configuration}
An electricity network models consists of nodal and series elements. Loads, generators, and shunts are nodal elements and are taken into account in the complex power vector $\mathbf{S}$. Transformers, line elements, and regulators are placed in series. They are accounted for in the $Y_{bus}$ matrix. Regulators are elements that are placed alone the distribution feeder to regulate the voltage profile of the feeder.  

@@

Loads are modeled as buses where active and reactive power ($P$ and $Q$) are specified. They are modeled as a function of the voltage in which the load-model type determines the relation between power and voltage. Three types exist: the constant power (P), constant current (I), and constant impedance model (Z). Together called: ZIP-load models.    

@@

    Constant power loads draw the same power from its source even if the source changes voltage.

@@

    A constant current load varies its internal resistance according to the voltage which is being fed to it in order to achieve a constant current despite fluctuations in the source voltage.  
    % \hl{quote: A constant current load is one which varies its internal resistance to achieve a constant current regardless of the voltage which is being fed to it and therefore the power will vary. In other words, a constant current load would draw approx the same amount of current, despite wide variations in the source voltage or other impedances in the circuit.}

@@

    A constant impedance load presents the same impedance even when voltage is fluctuating. 

@@

Loads are physically connected in a grounded Wye or ungrounded Delta configuration. In the grounded Wye configuration all the three phases are connected to a single neutral point. This neutral point is connected to the ground by a fourth wire: the neutral conductor. The voltage is specified line-to-neutral.
In an ungrounded Delta configuration the loads are connected phase-to-phase without a neutral conductor. Here, the voltage is specified phase-to-phase, also called line-to-line. The line-to-line voltages of a three-phase Delta-configuration are: $V^{ab}$, $V^{bc}$, and $V^{ca}$.  The line-to-line voltages and line-to-neutral voltages ($V^a$, $V^b$, and $V^c$) are related as follows:

@@

We solve the three-phase distribution systems with the NR-TCIM method, which implies that we need the specified currents $I_s$ in our mismatch equation (eq: \ref{eq:curmism}) instead of specified powers. As currents are never measured in a power system, they are conducted from the specified power according the relationship: 

@@

If we divide the expressions of the loads $S$ by $V$ and take the conjugate of this new relation we get the load models expressions in terms of specified current:

@@

\subparagraph{Wye - Z }
\begin{equation}
    I := (\frac{S|V|^2}{V})^* = ({P_0+\iota Q_0})^*\frac{|V|^2}{V^*} = ({P_0+\iota Q_0})^*\frac{VV^*}{V^*} =  ({P_0+\iota Q_0})^*V
\end{equation}
Or in three-phase matrix: 
\begin{align}
\begin{bmatrix}
    I^a\\
    I^b\\
    I^c
    \end{bmatrix}_i = \begin{bmatrix}
   ({ P^a +\iota Q^a})^*{V^a}\\
    ({ P^b +\iota Q^b})^*{V^b}\\
    ({ P^c +\iota Q^c})^*{V^c}
    \end{bmatrix}_i 
    \end{align}
\paragraph{Delta loads}
Delta loads are expressed by line-to-line voltages and currents while we need line-to-neutral currents in our calculations. We express the line-to-neutral currents directly for the three load models using \begin{align}
    I^a=I^{ab}-I^{ca},\\
    I^b=I^{bc}-I^{ab},\\
    I^c=I^{ca}-I^{bc}.
\end{align}
\subparagraph{Delta - P}
\begin{align}
\begin{bmatrix}
    I^a\\
    I^b\\
    I^c
    \end{bmatrix}_i := \begin{bmatrix}
    I^{ab}-I^{ca}\\
    I^{bc}-I^{ab}\\
    I^{ca}-I^{bc}
    \end{bmatrix}_i = \begin{bmatrix}
   (\frac{ P^{ab} +\iota Q^{ab}}{V^a-V^b})^* - (\frac{ P^{ca} +\iota Q^{ca}}{V^c-V^a})^* \\
    (\frac{ P^{bc} +\iota Q^{bc}}{V^b-V^c})^* - (\frac{ P^{ab} +\iota Q^{ab}}{V^a-V^b})^*\\
    (\frac{ P^{ca} +\iota Q^{ca}}{V^c-V^a})^* - (\frac{ P^{bc} +\iota Q^{bc}}{V^b-V^c})^*
    \end{bmatrix}_i 
    \end{align}
\subparagraph{Delta - I }
\begin{align}
\begin{bmatrix}
    I^a\\
    I^b\\
    I^c
    \end{bmatrix}_i = \begin{bmatrix}
    ({ P^{ab} +\iota Q^{ab}})^*\frac{(V^a-V^b)}{|V^a-V^b|} - ({ P^{ca} +\iota Q^{ca}})^*\frac{(V^c-V^a)}{|V^c-V^a|}  \\
    ({ P^{bc} +\iota Q^{bc}})^*\frac{(V^b-V^c)}{|V^b-V^c|} - ({ P^{ab} +\iota Q^{ab}})^*\frac{(V^a-V^b)}{|V^a-V^b|}\\
    ({ P^{ca} +\iota Q^{ca}})^*\frac{(V^c-V^a)}{|V^c-V^a|} -  ({ P^{bc} +\iota Q^{bc}})^*\frac{(V^b-V^c)}{|V^b-V^c|}
    \end{bmatrix}_i 
    \end{align}

@@

\subsubsection{Transformers}
Transformers are placed between two networks that have different voltage levels. The primary side ($p$) of the transformer is connected to the higher voltage network and the secondary side ($s$) is connected to the lower voltage level. The transformer in a power system is represented by two blocks:
\begin{itemize}
    \item An admittance matrix block, $Y^{abc}_T$, consisting of per-unit leakage admittance $y_t$, and  
    \item A shunt block that models core losses. The shunt block can be represented by voltage dependent loads and follow the representation in Tab. \ref{tab:ZIPmodels}. 
\end{itemize}
The admittance matrix block is given in a similar manner as the $Y_{bus}$ matrix of a distribution cable, but now consisting of leakage admittance of the transformer: \begin{equation}
    Y^{abc}_T = \begin{bmatrix}
    Y_{pp}^{abc} & Y^{abc}_{ps} \\
     Y_{sp}^{abc} & Y^{abc}_{ss}
    \end{bmatrix}.
\end{equation}
Each block $Y^{abc}_{ij}$, $i,j\in\{p,s\}$ has a different configuration. This configuration depends on the connection of the transformer and is either one of the three following blocks: \begin{equation}\resizebox{0.91\columnwidth}{!}{%
$
    Y_1=\begin{bmatrix}
    y_t & 0 & 0 \\
    0 & y_t & 0 \\
    0 & 0 & y_t
    \end{bmatrix},\quad Y_2=\frac{1}{3}\begin{bmatrix}
    2y_t & -y_t & -y_t \\
    -y_t & 2y_t & -y_t \\
    -y_t & -y_t & 2y_t
    \end{bmatrix},\quad\mbox{and}\quad Y_3=\frac{1}{\sqrt{3}}\begin{bmatrix}
    -y_t & y_t & 0 \\
    0 & -y_t & y_t \\
    y_t & 0 & -y_t 
    \end{bmatrix}. $%
    }
\end{equation}
Tab. \ref{tab:trafoblocks} represents the corresponding block per transformer configuration. \\
The pu leakage admittance $y_t$ consists of resistance $r$ and reactance $x$.  The equations should be scaled according the correct base quantities in order to match the different pu-systems. The pu leakage admittance results in: 
\begin{equation}
    y_t = \frac{1}{z_t},\quad\mbox{where } z_t = (r+\iota x)(\frac{V_{low}}{V_{base}})^2\frac{S_{base}}{S_t},
\end{equation}
where $r$ and $x$ are given in fractions. $V_{low}$ is the voltage at the secondary level of the transformer, and $S_T$ is apparent power of the transformer, expressed in volt-ampere. $V_{base}$ is always given in line-to-line and should be converted to line-to-neutral by dividing it by $\sqrt{3}$, in order to match the calculations. 
% Several transformer connections exist, which define the choice of the four blocks of the admittance matrix $Y_T^{abc}$. The correct blocks are described in Tab. \ref{tab:trafoblocks}.

@@

Step-Voltage Regulators (SVRs) are installed along distribution feeders, often after the substation, but also at other locations, to control the voltage along the feeder and to keep it within an acceptable range \footnote{
In short, the SVR is a sequence of automatic transformers that can change their tap ratio. The tap ratio of a transformer is responsible for the voltage level at the secondary side of a transformer, the output voltage. In a SVR, this voltage is measured; if it falls outside the safety range, the tap ratio changes to a different position to adjust the output voltage. This continues until the voltage is within acceptable limits.
In our load-flow  computations, we are not implementing a SVR that changes its tap ratio automatically. We pre-determine the `safe' tap-ratios based on output data of openDSS. With these data we model the SVR directly}. \\
A SVR is modeled as a series block that is placed between any connected pair of buses ($n$,$m$). The  voltages and currents of the primary (bus $n$) and secondary side (bus $m$) of the SVR are related via matrices:  $A_V$, $A_I$, and $Z_R\in\mathbb{C}^{6x6}$. The first one represents voltage gain, the second current gain, and the last impedance. $A_I$ and $A_V$ consists of effective regulator ratio $a_R$ and $Z_R$ consists of pu series impedance $z_R$ of the SVR. $a_R$ is expressed as:

@@

    a_R = 1\mp 0.000625tap, 
\end{equation} where \textit{tap} is the tap-ratio of the SVR, that needs be determined upfront. 
Three-phase SVRs are commonly configured as Wye, closed-delta, and open-delta connections. Just like transformers, these configurations determine the entities of the three blocks $A_I$, $A_V$, and $Z_R$. Furthermore, it holds that:  \begin{equation}

@@

These three blocks form the SVR-admittance matrix, $Y^{abc}_{R}$. The admittance matrix $Y^{abc}_R$ looks as follows:

@@

A_I F_R^{-1}Y^{abc}_{nn}A_I^T & -A_I F_R^{-1}Y^{abc}_{nm} \\[7pt]

@@

$Y^{abc}_{ij},\ i,j\in\{n,m\}$ is the admittance of the distribution line along which the regulator is placed. $F_R$ and $F_R^{-T}$ are defined as follows: 

@@

    F_R &= \mathbf{I}_{3\times 3}+Y^{abc}_{nn}A_I^TZ_RA_I,\\[3pt]

@@

\paragraph{Ideal SVRs}
If the auto-transformer of the SVR is ideal, the tap-ratio is equal to $0$ and the impedance matrix $Z_R$ is equal $0$. This results into $F_R$ being the identity matrix: $\mathbf{I}_{3\times 3}$. The admittance matrix of an ideal SVR becomes the following:

@@

A_I Y^{abc}_{nn}A_I^T & -A_I  Y^{abc}_{nm} \\[5pt]

@@

We refer to paper \todo{source} on how to derive the blocks $A_I$, $A_V$, and $Z_R$. We list the entities of these blocks, per SVR configuration, in Tab. \ref{tab:svrent}.

@@

We have a single-phase transmission system described by the following power flow relation: 
 \begin{equation}\label{eq:SVIa}
   S_i^a=V_i^a\overline{I^a_i}=V_i^a\sum_{k=1}^N\overline{Y^a_{ik}}\overline{V^a_k}     ,
and a three-phase distribution system, described by the following power flow relation:
\begin{equation}\label{eq:SVIp}
    S_i^p=V_i^p\overline{I^p_i}=V_i^p\sum_{k=1}^N\sum_{q=a,b,c}\overline{Y^{pq}_{ik}}\overline{V^q_k},\quad p\in\{a,b,c\}.
These two networks are connected to each other via a substation. A dimension mismatch takes place at this substation, depicted in figure \ref{fig:boundary}. Table \ref{tab:substationbus} shows the single-phase and three-phase representation of $S$, $V$ and $Y$ in the network models. 
\begin{figure}[b]

@@

The literature suggests two approaches to run computations on integrated networks: a unified and a splitting approach, and it suggests two ways of modeling integrated networks: as a homogeneous or as a hybrid network. Two integrated three-phase networks form a homogeneous network\footnote{two single-phase networks can also make a homogeneous network, but we regard only unbalanced three-phase distribution networks.}: both the transmission and distribution network are modeled in three phases. A hybrid network consists of a single-phase transmission network and a three-phase distribution network. The unified methods first integrates the two network into one system and then solves the entire system at once. The splitting approach appoints the transmission network as the manager and the distribution network as the fellow. It iterates between the two networks and at each iteration, it solves the networks separately. We call the unified approach applied to homogeneous networks the full three-phase (F3P) method and applied to hybrid networks the interconnected (IC) method. We call all splitting methods manager-fellow splitting (MFS) methods. They are described in the literature  as Master-Slave splitting methods\cite{Sun2008}, but we change the naming convention. We define the MFS-methods based on the network model they are applied to, eg: the splitting approach applied to hybrid networks is called the MFS-hybrid method. Figure \ref{fig:algorithms} gives an overview of the methods. 

@@

\begin{figure}[h]
    \includegraphics[width=0.8\textwidth]{Images/substationtransformer.eps}
